---
title: "SABD Project: Urban Mobility Analysis"
subtitle: "Part 2 - Data Handling with R and MongoDB"
author: "I. Joly, C Bobineau, B. Agard"
date: "juin 2022"
output: pdf_document
header-includes:
  \renewcommand{\contentsname}{Table of Content}
  \usepackage{tcolorbox}
  \newtcolorbox{blackbox}{
  colback=black,
  colframe=orange,
  coltext=white,
  boxsep=5pt,
  arc=4pt}
toc: true
theme: unitedtlm
link-citations: yes
biblio-style: apalike
bibliography: ["biblio/Project_biblio.bib"]
nocite: |
  @xie2015, @Chang2013, @xie2016, @Munzner2015, @harrell2013regression
---

<!-- Installation Instructions  -->
```{r , eval=F, echo=FALSE}
# This Rmd document is in the main folder
# Project should habe subfolder : "biblio" with "Project_biblio.bib" file
# data file "allgre.PB_V2.RData" is in the main 
# The FOND_MAP folder contain all the NewESRI.* files
#    They are usefull to test GIS representation
```

```{r definitions, echo=FALSE, results='hide'}
duedate <- "20/01/2023"

# In the Part2_GIS.Rmd file you will find the  Parameter map_draw  to turn TRUE to draw maps (or FALSE)
```




```{r setup, include=FALSE, cache=FALSE}
## Libraries
## You may have to install these libraries first 
## with the tab on bottom right panel : "Packages" ; click install
## or : install.packages("ggplot2")   - don't forget the ""
library(tidyverse)
library(knitr)
library(ggplot2)
library(skimr)
#install.packages("Mongolite")
library(mongolite)

## Global options
knitr::opts_chunk$set(               # Here we define behaviours of the code chunks in the Rmd
               echo=T,        # to display in the generated document code lines (False to hide them; but they will be ran)
               eval=FALSE,     # to run the chunks
	             message=FALSE, # to hide in the document R messages while running chuncks
               warning=FALSE) # to hide in the document R alerts
knitr::opts_knit$set(width=75)  
options(max.print="75")       # to define max number of elements (lines) displayed by R when running an instruction
```


# Statement

## Objectives of the Part 2

Part 1 ends with saving the data. You should know how to create the reproducible document, load and save the data for the Mobility project.

Part 2 aims at guiding you to handle the data: select the variables, modify and create variables, either with R or Mongo or both.

First, the document presents several operations applied to the variables, to create new ones, factorize some, etc.

This part ends with some operations in the objective to create GIS map based on the data: display a map of Grenoble metropolitan area with color indicating some variables values (average daily travel time and number of trip)

You will continue with data handling for your own topic.

First lets load the data

```{r eval=TRUE}
load("allgre.PB_V2.RData")
```

<!-- Variable handling with R   -->
```{r eval=TRUE, child='Part2_Variables.Rmd'}
```

\newpage

<!---   You will find in this child Rmd document the parameter: map_draw to turn TRUE or FALSE -->

<!-- GIS Variable handling with R   -->
```{r eval=TRUE, child='Part2_GIS_2023.Rmd'}
```





\newpage

# References

<div id="refs"></div>



## Sources

(https://www.mongodb.com/languages/mongodb-and-r-example)

(https://rtavenar.github.io/mongo_book/content/07_scripts_R.html)

(https://jeroen.github.io/mongolite/query-data.html#sort-and-limit)

\newpage

# Appendix