
:::: {.blackbox data-latex=""}
In this part, we work with geographical data to produce maps

Instructions in chunks transform the data and perform some calculus for GIS maps, they should be reviewed in Mongo

Note that the functions for map creation maybe time and resource consuming ! That is why the chunk are `eval=FALSE`: you will have to turn them on if you want to draw the map.

::::




```{r}
#Parameter to turn TRUE to draw maps (or FALSE)
map_draw <- TRUE
```
  
  
# GIS Part

```{r}
#install.packages("sf")
#install.packages("terra")
#install.packages("spData")
#install.packages("spDataLarge", repos = "https://nowosad.r-universe.dev")

library(sf) 
library(cartography)
library(terra)  
library(spData)        # load geographic data
library(spDataLarge) 
library(RColorBrewer)
newESRI <- st_read("FOND_MAP/newESRI.shp")
#st_read("G:/Ira_Lessons_with_Rmd/2021_SABD_SmartAnalytics/2022_TP_EMD/FILES_OK/FOND_MAP/newESRI.shp")

choroLayer(
  x = newESRI, var = "cod_sectir", 
  legend.horiz = FALSE, 
  legend.pos = "topright",
  legend.title.txt = "My first map : cod_sectir",
  legend.values.rnd = -2
)
```




# Map and Calculus of area indicators: CARTE NBD BTT

:::: {.blackbox data-latex=""}
Here 
\begin{itemize}
\item Calculation by  zone: average and median of the daily travel time on individuals in the area and the number of trips and total number of individuals in the area
\item Reduction of the DF to lighten the execution
\item Merging of the DF with the GIS DF (which contains points coordinates and polygons)
\end{itemize}
::::
  
```{r,  results='hide', message=FALSE, eval= map_draw}
DF <- allgreI[, c("nbd", "btt", "tir", "zoneres.y", "id_pers")]
# Calcul base individus des indicateurs individuels
DF_Indiv <- na.omit(DF)
addmargins(table(allgreI$zoneres.y))
DF_Indiv$BTT_moy <- ave(DF_Indiv$btt,DF_Indiv$tir, FUN = mean)
DF_Indiv$BTT_med <- ave(na.omit(DF_Indiv$btt),DF_Indiv$tir, FUN = median)
DF_Indiv$nbd_med <- ave(DF_Indiv$nbd,DF_Indiv$tir, FUN = median)
DF_Indiv$nbd_moy <- ave(DF_Indiv$nbd,DF_Indiv$tir, FUN = mean)
DF_Indiv$Maxind <- ave(DF_Indiv$id_pers,DF_Indiv$tir, FUN = max)
# Réduction à la base Zone (avec le max(ind) par TIR)
DF_Zone <- subset(DF_Indiv, DF_Indiv$id_pers == DF_Indiv$Maxind, select =c(zoneres.y,BTT_med, BTT_moy, nbd_moy, nbd_med,tir))
``` 

### Merging with newESRI and data

```{r eval=map_draw}
# library(sp) # the trick is that this package must be loaded!
# MERGING with option duplicateGEOM=T to get as many polygons as the number of observationsin the data
DF_Zone$cod_sectir <- DF_Zone$tir
newESRI_DF <- merge(newESRI, DF_Zone, by = "cod_sectir", duplicateGeoms = TRUE)
```

### Projecting your geographic coordinates to the equirectangular projection

After merging the data and the map, geographic coordinates have to be projected.

https://r-spatial.org/r/2020/06/17/s2.html


```{r}
newESRI_DF = st_transform(newESRI_DF, "+proj=eqc")
```




# Colored mapping 
```{r,  message=FALSE, eval=map_draw}
choroLayer(spdf = newESRI_DF, var = "tir")
title("Mapping of the area number - not really useful")

table(newESRI_DF$tir)
```

# Map with labels 

It may need to install `lwgeom`

```{r eval=map_draw}

choroLayer(spdf = newESRI_DF,var = "tir")
# Label plot of the Mean
labelLayer(spdf = newESRI_DF, txt = "tir", col = "#690409", cex = 0.9, font = 2)
```



# Complete map of daily travel times

The script may need to be run out of  MARKDOWN

```{r, echo=TRUE, eval=map_draw}
Nclasse <- 5
########## Carte colorée
choroLayer(spdf = newESRI_DF, 
           #df = DF_Zone,
           spdfid = "cod_sectir", # id in map data
           dfid = "tir",  # id in dataframe
           var = "BTT_moy", 
           nclass = Nclasse,
           col = rev(brewer.pal(n = Nclasse, "RdYlBu")),
           legend.pos = "right", legend.frame = TRUE,
           legend.title.txt = "BTT moyens par zone")
########### Ajout de label : nbd par zone
# Label creation 
newESRI_DF$lab <- paste(round(newESRI_DF$nbd_moy,1), sep ="")
# Label plot of the Mean
labelLayer(spdf = newESRI_DF,  txt = "lab", col = "#690409", cex = 0.9, font = 2)
####### Ajout d'éléments de présentations : Layout plot
layoutLayer(title = "BTT moyens par zone",
            sources = "EMD Grenoble 2010",
            author = "I. Joly",
            scale = 0,
            north = T,
            frame = TRUE,
            col = "black",
            coltitle = "white")
```
