


## Part Two - Data processing


Here follow multiple data handling operations. 


The following script elements reduce the database to trip, person and household level.

They also give some guidance on how to aggregate the information available at a lower level.

They are written using R base functions and tidyverse functions. Just choose the one you want to apply.

### Number of trips

Analyzing the number of trips, you should check that 44 observations are: `nbd = 22`

These are persons that declare 22 trips in their day. For one individual this value is repeated 22 times in the DF ! We deduce that on ly 2 persons have done 22 trips in their day...

```{r echo=TRUE, eval=T}
table(allgre.PB_V2$nbd)
```


```{r echo=TRUE, eval=T}
# or
DF <- allgre.PB_V2 %>% dplyr::select(sexe, nbd) %>% table()
kable(DF)  
```

`nbd` was determined counting the trips in the DF for each individual.  `NO_DEPL` is the number of trips determined by the survey institution.

Studying the number of trips (`NO_DEPL`), we see that 2 trips are numbered 22.

```{r echo=TRUE, eval=T}
# Check of NO_DEPL - number of trips
table(allgre.PB_V2$NO_DEPL)
```

We can check that each trip (observation) is associated to an unique  `id`.

We see that `r table(allgre.PB_V2$NO_DEPL)[2]` trips have no number. These are lines for individuals without any trips: immobile persons

> A half empty line describe the individual but not the trip


Identification key is created for each trip using individual ID and the trip number

```{r echo=TRUE, eval=T}
# Creation of new id_depl
allgre.PB_V2$id_depl <- allgre.PB_V2$id_pers * 10  + allgre.PB_V2$NO_DEPL 
table(is.na(allgre.PB_V2$id_depl))
table(allgre.PB_V2$nbd)
```

Or

```{r}
allgre.PB_V2 %>% dplyr::mutate(
  id_depl = id_pers * 10  + NO_DEPL) %>%
  select(id_depl) %>% is.na() %>%   table()
allgre.PB_V2 %>% select(nbd) %>% table()
```

\newpage

## New Variables Definition

### Immobil person

+ we point individual that have at least one trip in the day (`nbd != O` $\leftrightarrow nbd \neq 0$)

```{r}
allgre.PB_V2$UN <- as.numeric(allgre.PB_V2$nbd !=0)
table(as.numeric(allgre.PB_V2$UN))
```

or

```{r, message=FALSE}
allgre.PB_V2 <- allgre.PB_V2 %>% dplyr::mutate(  UN =  as.numeric(nbd !=0) )

allgre.PB_V2 %>%  dplyr::select(UN)  %>% table()
```

### Area number redefinition

+ Location area of origin and destination of trips are recoded through a simple transformation:


```{r, message=FALSE}
allgre.PB_V2 <- allgre.PB_V2 %>% dplyr::mutate(  
  UN =  as.numeric(nbd !=0) ,
  DEST = round(zonedest/1000),
  ORIG = round(zoneorig/1000) ,
  )

head(allgre.PB_V2$DEST)
head(allgre.PB_V2$ORIG)
table(allgre.PB_V2$ORIG)
```




### Average travelled distances per area

+ we calculate the average distance as crow flies (`D12`) and the distance through the transport network (`D13`)
+ we count the number of `NA` (non observed) distances (`D12`) 
+ we calculate the average travelled distance (`D12`) given the residential location of individual

> averages are calculated at the residential location area; but non stored in the DF!

```{r}
mean(allgre.PB_V2$D12, na.rm = T)
mean(allgre.PB_V2$D13, na.rm = T)
summary(allgre.PB_V2$D12)
# nombre de distance 'NA' par zone
table(allgre.PB_V2$zoneres.y[is.na(allgre.PB_V2$D12)==1])
# tapply(allgre.PB_V2$D12, as.factor(allgre.PB_V2$zoneres.y),mean, na.rm = TRUE)
```

Or

```{r eval=T}
# allgre.PB_V2 %>% select(D12, D13) %>% skim()

allgre.PB_V2 %>% select(D12, D13) %>% summarize_each(funs(mean(., na.rm = TRUE)))

allgre.PB_V2 %>% 
  group_by(zoneres.y) %>% 
  select(D12, D13) %>% 
  summarize_each( funs(mean(., na.rm = TRUE)))

allgre.PB_V2 %>%
  group_by(zoneres.y) %>%
  summarise(across(D12:D13,~mean(.x, na.rm = TRUE)))
```

## Factorisation

+ We define qualitative variables (categorical variables) as **factor** and we associate labels to their modalities

```{r}
## GRENOBLE 2010		 :
#######
table(allgre.PB_V2$OCCU1 )
allgre.PB_V2$statut2 <- factor(allgre.PB_V2$OCCU1 , 
                               labels=c("TRAVAIL A TEMPS PLEIN",
                        "TRAVAIL A TEMPS PARTIEL","FORMATION STAGE","ETUDIANT", "SCOLAIRE",
                        "CHOMEUR","RETRAITE", "RESTE AU FOYER", "AUTRE"))
table(allgre.PB_V2$statut2)
#####
levels(allgre.PB_V2$statut2) <-  c("ACTIF","ACTIF", "ACTIF", "ETUDIANT",
                                   "SCOLAIRE", "CHOMEUR","RETRAITE", 
                                   "RESTE AU FOYER", "AUTRE")
table(allgre.PB_V2$statut2)
#####
allgre.PB_V2$cspgroup <- trunc(allgre.PB_V2$csp / 10)
allgre.PB_V2$cspgroup[allgre.PB_V2$pcs %in% c(81,82)] <- 7
allgre.PB_V2$cspgroup[is.na(allgre.PB_V2$cspgroup)==T] <- 9
table(allgre.PB_V2$cspgroup)


allgre.PB_V2 <- allgre.PB_V2 %>% mutate(
  mot_o_red = factor( case_when(
                motifor %in% c( 1, 2)	~ 1, #  domicile & residence secondaire
                motifor %in% c(11, 12) 	~ 2, 	#	/*travail*/
                motifor %in% c(21, 22, 26)	~ 3, 	#	/*primaire*/
                motifor %in% c(23, 24, 27, 28)  ~ 4 ,	#	/*col-lyc*/
                motifor %in% c( 25, 29)		~ 5,  # /*univ*/
                30 < motifor & motifor < 44	~ 6,	#	/*achat serv*/
                 49 < motifor & motifor < 55	~ 7,	#	/*loisir*/
                 60 < motifor & motifor < 75	~ 8,	#	/*acct*/
                 motifor == 91 | motifor == NA	~ 9, #	/*autre*/
                 motifor == 81 		~ 10	,  #/*tournee prof*/
                (motifdes == 91 | motifdes == NA	) & nbd==0 ~ 0
                # else 0
                ))
  , mot_d_red = factor( case_when(
                motifdes %in% c( 1, 2)	~ 1, #  domicile & residence secondaire
                motifdes %in% c(11, 12) 	~ 2, 	#	/*travail*/
                motifdes %in% c(21, 22, 26)	~ 3, 	#	/*primaire*/
                motifdes %in% c(23, 24, 27, 28)  ~ 4 ,	#	/*col-lyc*/
                motifdes %in% c( 25, 29)		~ 5,  # /*univ*/
                30 < motifdes & motifdes < 44	~ 6,	#	/*achat serv*/
                 49 < motifdes & motifdes < 55	~ 7,	#	/*loisir*/
                 60 < motifdes & motifdes < 75	~ 8,	#	/*acct*/
                 motifdes == 91 | motifdes == NA	~ 9, #	/*autre*/
                 motifdes == 81 		~ 10	,  #/*tournee prof*/
                (motifdes == 91 | motifdes == NA	) & nbd==0 ~ 0
                # else 0
                ))
  )

motif.red <- c("DOMICILE" , "TRAVAIL" , "ECOLE PRIMAIRE" , "C.E.S.-LYCEE" ,"UNIVERSITE" ,'ACHAT' ,'LOISIR' ,'ACCOMPAGNEMENT' , "AUTRES", "Professionnel")
allgre.PB_V2$mot_o_red <- factor(allgre.PB_V2$mot_o_red,  labels =  motif.red)
allgre.PB_V2$mot_d_red <- factor(allgre.PB_V2$mot_d_red,  labels =  motif.red)

```

> Explain what is done in the preceeding operations

### Some variables definitions

For many variables we calculate a sum, a max or mean over a specific group of observations: for ex, the individual sum of trips duration 


In these case a strategy maybe to calculate in the DF of trips the indicators and then reduce the DF to the individual level.

> Below, we define the variables at the relevant level and then reduce the DF at the relevant level for anlysis

## Stops DataFrame : 

+ We count the number of stays in a trip. 

```{r}
#    BASE DEPLACEMENTS
## exemple de calcul : nombre de trajets par déplacement
allgre.PB_V2$nbt2 <- ave( allgre.PB_V2$NO_TRAJ, allgre.PB_V2$id_depl, FUN = max)
addmargins( table(allgre.PB_V2$nbt2) )
```


### Trip DataFrame: *Déplacements*

+ Total number of stays of each trip of each person
+ Reduction to the DF where 1 line = 1 trip of one person
+ we keep the trips ID without double
+ then we exclude immobile persons
+ finally, we perform some counts and calculus on number of trips, given the transport mode, the trip duration
+ the DF is reduced to the trip DF level

```{r}
## Réduction au déplacement : ATTENTION On ejecte les immobiles
allgreD <- distinct(allgre.PB_V2, id_depl, .keep_all = T)
```

### Person DataFrame *Individus*


+ Total travel time in a day of each person
+ total number of trips in a day of each person

```{r}
############
#    BASE INDIVIDUS
## exemple de calcul : durée quotidienne des déplacements d'un individu
allgreD$btt2 <- ave( allgreD$duree, allgreD$id_pers, FUN = sum)
cpte <- function(x){length(unique(x)) }
allgreD$nbd2 <- ave( allgreD$NO_DEPL, allgreD$id_pers, FUN = cpte )
```

+ Reduction to  the person level DF using the individual ID

```{r}
## Reduction
allgreI <- distinct(allgreD, id_pers, .keep_all = T)
```

```{r}
summary(allgreI$btt); summary(allgreI$btt2)
addmargins( table(allgreI$nbd) )
table(is.na(allgreI$nbd))
addmargins( table(allgreI$nbd2) )

# nombre d'incohérences entre Nbd Nbd2
sum(allgreI$nbd != allgreI$nbd2)
```


Remark: to calculate the daily travel time (BTT), elements `btt` and `btt2` differ due to preprocessing operations to calculate `btt` (correction of error in the declared time of trip)


+ Number of person in a household, as the maximum of the person number in a household
+ We determine if a least one person in the household holds a driver licence  (using the maximum of the "permis de conduire" variable in the household)


```{r}
############
#    BASE MENAGE
## exemple de calcul : taille du ménage et permis-ménage
allgreI$taillemng <- ave(allgreI$NO_PERS, allgreI$id_men, FUN = max)
allgreI$Permis_mng <- ave(allgreI$permis, allgreI$id_men, FUN = max)
addmargins( table(allgreI$taillemng) )
```



+ Reduction to the household level DF, based on the household ID

```{r}
## Reduction
allgreM <- distinct(allgreI, id_men, .keep_all = T)
    #subset(allgreI, allgreI$NO_PERS == allgreI$taillemng )
addmargins( table(allgreM$taillemng) )
addmargins( table(allgreM$Permis_mng) )

table(allgreI$freqtcu)
table(allgreI$nbd)
table(allgreI$freqter)
```

## External Calculus

+ Modal share in the DF

```{r}
# Par mode
table(allgreD$mode_depl_ag)
```

+ Trip duration

```{r}
# durée
addmargins( table(allgreD$duree))

table(is.na(allgreD$duree))
table(cut(allgreD$duree, breaks=seq(0, max(allgreD$duree, na.rm = T), 30)), useNA = "always")
```

## Save of the 3 DataFrames: Trip, Person, Household

```{r eval=FALSE, echo=TRUE} 
#ici on a r eval= FALSE 
allgreD <- allgreD[, c( c(1:4), c(75:105), 106:109, 112, 114:116, 117:119)]
allgreI <- allgreI[, c( c(1:3), c(46:74) , 106, 107, 111, 113, 117, 120, 121 )]
allgreM <- allgreM[,c( c(1,2), c(5:45), 106:109, 110, 124,125  )]
save(allgreD, file = "allgreD.RData")
save(allgreI, file = "allgreI.RData")
save(allgreM, file = "allgreM.RData")
```


## **Your** Data processing

**You should start here**
You can download these data files and load them:

```{r Load_Data, eval=TRUE}
load("allgreD.RData")
load("allgreI.RData")
load("allgreM.RData")
```
